\plaintitle

### 2017-01-07 drafts/labor/piecework_lit.tex
piecework review refactor

\subsection{Case studies in piecework}
\topic{Throughout the rest of the paper, we will return to three major case studies to frame our analyses:
railroad and other industrial workers;
Airy's employment of \textit{human computers};
and domestic and farm work (specifically, the ``match--girls'').}
In introducing these cases at a high level,
we'll trace the history of piecework
while also framing the later analysis of the major research threads we named earlier
(questions toward complexity, decomposition, and relationships, respectively).

@@

\subsubsection{\ali{should I drop this? Railroad and other }Industrial Workers}
\topic{Piecework might be most familiar to the HCI researcher in the context of the assembly line, which largely defined manufacturing through the \nth{20} century.}
In the railroad industry and especially on the assembly line
we see many of the mechanisms of work and worker management that later made their way into crowd work.
Furthermore, it's from here the piecework literature draws most upon to find limits speaking to complexity and task granularity.
We'll expose these facets of piecework very superficially in the context of industrial work
to better speak to these research questions in crowd work later.

\topic{Railway companies adopted piecework regimes in the early \nth{20} century;
what followed was a flourishing in the management practices with pieceworkers,
as railway companies struggled to find effective ways
to motivate and evaluate this skilled workforce of engineers.}
\citeauthor{10.2307/23702539} takes up a case study of the Santa Fe Railway,
finding that they employed ``efficiency experts'' to develop a ``standard time''
to determine pay for each task at the company informed by
``thousands of individual operations''; % ~\cite{10.2307/23702539}.
\citeauthor{10.2307/23702539} goes on to list
some of the roles required to facilitate piecework
in the early \nth{20} century
--- among them, ``piecework clerks, inspectors, and `experts'~''~\cite{10.2307/23702539}.

\topic{Piecework through the \nth{20} century centered around auto and other mass manufacturing, but
found its way into the war effort during World War II.}
With the vast majority of men drafted into service,
factories found themselves turning to
a predominantly female workforce that had neither
the formal training nor
the years of apprenticeship experience
that conventional workers would have had.
Rather than attempting to train this new labor force in every aspect of industrial work,
these women were trained for individual tasks
% --- such as riveting component to another ---
and assigned to that task. %, decomposed from a more comprehensive task.
One might reflect on the observation that ``Rosie the Riveter'',
an icon of \nth{20} century America who
represented empowerment and opportunity for women~\cite{honey1985creating},
was a pieceworker~\cite{davies2014origins}.

% \msb{even further backwards??? I would have expected this section to take a time slice and tell me about one group of people in detail}
% \msb{this content might go above in the paragraph about the heydey of piecework in the piecework review}
\topic{The 1930s represented a boom for piecework on an unprecedented scale,
especially among engineering and metalworking industries.}
\citeauthor{hart2013rise} characterize the 1930s
--- and more broadly the first half of the \nth{20} century ---
as the ``heyday'' of the use of piecework.
They attribute this to the shortage of male workers,
who would have gone through a conventional apprenticeship process
affording them more comprehensive knowledge of the total scope of work.

@@

\topic{The application of piecework to farm work in the late \nth{19} century and
later to manufacturing of small goods, such as garments and matches, at the turn of the \nth{20} century
proved to be a formative period for piecework as we would come to know it.}
Here, then, we'll discuss the implementation of piecework regimes in farms,
in homes where workers assembled clothing and were paid by the \textit{piece} of output, and
finally in the early sweatshops where women made matchsticks under dangerous and even hazardous conditions.

\topic{Farm and field work emerged early on as a rich field for the use of piecework,
perhaps in part due to the relatively low skillset necessary to complete work.}
We discussed earlier that farm work applied the idea of piecework to the farm by
paying workers for things like picking bushels of fruit or bringing to birth animals~\cite{10.2307/2338394}.
\citeauthor{hughRaynbirdTaskWork} argues from another perspective, but along a similar thread, that
the only factor important in piecework was
whether the worker's final product was to the ``satisfaction of his master''~\cite{hughRaynbirdTaskWork}.
Piecework's advantage,
\citeauthor{hughRaynbirdTaskWork} argues,
was this hands--off management of workers.

\topic{Piecework was beginning to find its way into homes near the end of the \nth{19} century.}
Textile manufacturers found that they could deliver fabric to people at their homes,
paying them to sew together clothing.
The manufacturers would later return to retrieve the finished garments,
paying these workers for each \textit{piece} of clothing made.
They could, in principle, assemble as much or as little clothing as they wanted;
the reality was more grim, as
\citeauthor{riisOtherSideLives} documented in \citeyear{riisOtherSideLives}~\cite{riisOtherSideLives}.
He found that
workers endured bleak living conditions and
worked long hours attempting to scrape together a living.

\topic{Piecework played a role in early mass manufacturing during the Industrial Revolution,
as well as the early years of worker collective organization and action.}
At the end of the \nth{19} century,
manufacturers employed women to assemble matchsticks in factories.
These match--makers were some of the first workers in mass manufacturing
to successfully rally for political causes,
first in the form of a march on parliament in 1871 to protest a proposed tax, and 
later (more famously) in what was later called ``the match--girls strike of 1888''~\cite{10.2307/3827491}.

@@

\topic{To understand how these women became such influential, if underreported, figures in labor advocacy's history,
we should shine a light onto the reputation that had formed for these workers.}
Match girls were the only group in \citeyear{booth1903life} to have formed a trade union,
according to \citeauthor{booth1903life}'s account at the time~\cite{booth1903life}.
\citeauthor{10.2307/3827491} noted that match girls
``\dots~pooled their resources to purchase their plumes and clothes~\dots~and expressed their solidarity through small [and major] strikes''~\cite{booth1903life}.
But they were also, as \citeauthor{10.2307/3827491} confesses, known for ``brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
These were workers who treasured their independence, but also fiercely protected one another, contributing to the common good.
Their ``brashness'' for instance may have detracted from their public image, but almost undoubtedly contributed to their sense of solidarity,
making their propensity to act against such unfair treatment and poor conditions understandable and maybe predictable.

@@

@@

\msb{This one seems way thinner than the others? Maybe it should be merged with the earlier description of Airy.}
\topic{Some of the first systematic cases of what we would recognize as crowd work
can be found in the study of astronomy.}
In the \nth{19} century, the calculation of celestial bodies had become
a competitive field with 
Airy needed to compute tables that would
allow sailors to locate themselves by starlight from sea.
This work ostensibly called for educated people who comprehensively understood mathematics.
Airy realized that he could break the tasks down and delegate the constituent parts
to ``human computers'' who
``\dots~possessed the basic skills of mathematics,
including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
As a result, many of Airy's computers had relatively rudimentary educations
compared to the background of education that typically worked in the calculation of solar tables.
Airy distributed tasks by mail,
allowing work to be completed by a somewhat geographically distributed workforce,
and paid for each piece of work completed.
Airy also instituted a policy of firing his computers once they reached age 23
\ali{this has a ``gosh isn't that interesting?'' feeling to it.
should i just leave it out until later when there's something to say about it
(namely, that the mode of work was initially designed to stymie professional growth)?}%
.

\ali{This should go later someplace, but not clear where:

This practice ensured two outcomes that arguably disfavored workers.
First, it eliminated any potential to advance professionally, as
workers' careers in this area ended relatively early in their careers,
and without formal education in mathematics they struggled to find work for which their experience was meaningful.
Second, it limited workers' ability to organize
by ensuring that workers hardly spent sufficient time to successfully rally their peers.}

>>>>

\subsection{Case studies in piecework}
\topic{Throughout the rest of the paper, we will return to four case studies to frame our analyses:
Airy's employment of human computers;
domestic and farm work;
the ``match--girls'' strike;
and industrial and assembly-line work.}
In introducing these cases at a high level,
we'll trace the history of piecework
while also framing the later analysis of the major research threads we named earlier: complexity, decomposition, and relationships.

@@

\msb{This one seems way thinner than the others? Maybe it should be merged with the earlier description of Airy.}
\topic{Some of the first systematic cases of what we would recognize as crowd work
can be found in the study of astronomy.}
In the \nth{19} century, the calculation of celestial bodies had become
a competitive field with 
Airy needed to compute tables that would
allow sailors to locate themselves by starlight from sea.
This work ostensibly called for educated people who comprehensively understood mathematics.
Airy realized that he could break the tasks down and delegate the constituent parts
to ``human computers'' who
``\dots~possessed the basic skills of mathematics,
including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
As a result, many of Airy's computers had relatively rudimentary educations
compared to the background of education that typically worked in the calculation of solar tables.
Airy distributed tasks by mail,
allowing work to be completed by a somewhat geographically distributed workforce,
and paid for each piece of work completed.
Airy also instituted a policy of firing his computers once they reached age 23.
%\ali{this has a ``gosh isn't that interesting?'' feeling to it.
%should i just leave it out until later when there's something to say about it
%(namely, that the mode of work was initially designed to stymie professional growth)?}%

The human computers captured several aspects of task decomposition that would become common. First, the work was designed such that it could be done independently and without collaboration. This enabled geographically-distributed workers to complete the tasks. Second, the work was designed so that intermediate results could be quickly verifiable as correct: Airy would have two workers each do the calculation, and a another person compare their answers. Third, Airy identified ways to decompose the large task into narrowly-trainable subtasks.
%\ali{This should go later someplace, but not clear where:

This practice ensured two outcomes that arguably disfavored workers.
First, it eliminated any potential to advance professionally, as
workers' careers in this area ended relatively early in their careers,
and without formal education in mathematics they struggled to find work for which their experience was meaningful.
Second, it limited workers' ability to organize
by ensuring that workers hardly spent sufficient time to successfully rally their peers.

@@

\topic{The application of piecework to farm work in the late \nth{19} century and
later to manufacturing of small goods, such as garments and matches, at the turn of the \nth{20} century
proved to be a formative period for piecework as we would come to know it.}
Piecework regimes in farms and in homes engaged workers in assembling clothing, where they were paid by the \textit{piece} of output. %, and
%finally in the early sweatshops where women made matchsticks under dangerous and even hazardous conditions.
%\topic{Farm and field work emerged early on as a rich field for the use of piecework,
%perhaps in part due to the relatively low skillset necessary to complete work.}
Textile manufacturers found that they could deliver fabric to people at their homes,
paying them to sew together clothing.
The manufacturers would later return to retrieve the finished garments,
paying these workers for each \textit{piece} of clothing made. 
Farm work applied the idea of piecework by
paying workers for things like picking bushels of fruit or bringing to birth animals~\cite{10.2307/2338394}.
\msb{cutting this next part, it seems irrelevant to the paragraph:}
%\citeauthor{hughRaynbirdTaskWork} argues from another perspective, but along a similar thread, that
%the only factor important in piecework was
%whether the worker's final product was to the ``satisfaction of his master''~\cite{hughRaynbirdTaskWork}.
%Piecework's advantage,
%\citeauthor{hughRaynbirdTaskWork} argues,
%was this hands--off management of workers.

Workers could, in principle, assemble as much or as little clothing as they wanted;
the reality was more grim, as
\citeauthor{riisOtherSideLives} documented in ``\citetitle{riisOtherSideLives}'' in \citeyear{riisOtherSideLives}~\cite{riisOtherSideLives}.
He found that
workers endured bleak living conditions and
worked long hours attempting to scrape together a living.

\subsubsection{The Match--Girls' Strike}

%\topic{Piecework played a role in early mass manufacturing during the Industrial Revolution,
%as well as the early years of worker collective organization and action.}
\topic{Match--makers were some of the first workers in mass manufacturing
to successfully rally for political causes.}
At the end of the \nth{19} century,
manufacturers employed women to assemble matchsticks in factories.
These women rallied first in the form of a march on parliament in 1871 to protest a proposed tax, and 
later (more famously) in what was later called ``the match--girls strike of 1888''~\cite{10.2307/3827491}.

@@

%\topic{To understand how these women became such influential, if underreported, figures in labor advocacy's history, we should shine a light onto the reputation that had formed for these workers.}
Match girls were the only group in \citeyear{booth1903life} to have formed a trade union,
according to \citeauthor{booth1903life}'s account at the time~\cite{booth1903life}.
\citeauthor{10.2307/3827491} noted that match girls
``\dots~pooled their resources to purchase their plumes and clothes~\dots~and expressed their solidarity through small [and major] strikes''~\cite{booth1903life}.
But they were also, as \citeauthor{10.2307/3827491} confesses, known for ``brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
These were workers who treasured their independence, but also fiercely protected one another, contributing to the common good.
Their ``brashness'' for instance may have detracted from their public image, but almost undoubtedly contributed to their sense of solidarity,
making their propensity to act against such unfair treatment and poor conditions understandable and maybe predictable.

@@

\subsubsection{Industrial Workers}
\topic{Piecework might be most familiar to the HCI researcher in the context of industrial and factory work, which largely defined manufacturing through the \nth{20} century.}
%In the railroad industry and especially on the assembly line we see many of the mechanisms of work and worker management that later made their way into crowd work.
%Furthermore, it's from here the piecework literature draws most upon to find limits speaking to complexity and task granularity.
%We'll expose these facets of piecework very superficially in the context of industrial work to better speak to these research questions in crowd work later.
\topic{Before the factory assembly line arose, railway companies adopted piecework regimes in the early \nth{20} century. \msb{What was pieceworky about this? They were paid per meter of railroad that they laid down?}
What followed was a flourishing of piecework management practices,
as railway companies worked to find effective ways
to motivate and evaluate this skilled workforce of engineers.}
\citeauthor{10.2307/23702539} takes up a case study of the Santa Fe Railway,
finding that they employed ``efficiency experts'' to develop a ``standard time''
to determine pay for each task at the company informed by
``thousands of individual operations''; % ~\cite{10.2307/23702539}.
\citeauthor{10.2307/23702539} goes on to list
some of the roles required to facilitate piecework
in the early \nth{20} century
--- among them, ``piecework clerks, inspectors, and `experts'~''~\cite{10.2307/23702539}.

\topic{The 1930s represented a boom for piecework on an unprecedented scale,
especially among engineering and metalworking industries.}
\citeauthor{hart2013rise} characterize the 1930s
--- and more broadly the first half of the \nth{20} century ---
as the ``heyday'' of the use of piecework.
They attribute this to the shortage of male workers,
who would have gone through a conventional apprenticeship process
affording them more comprehensive knowledge of the total scope of work.

\topic{Piecework %through the \nth{20} century centered around automobile and other mass manufacturing, but
found its way into the war effort during World War II.}
With the vast majority of men drafted into service,
factories found themselves turning to
a predominantly female workforce that had neither
the formal training nor
the years of apprenticeship experience
that conventional workers would have had.
Rather than attempting to train this new labor force in every aspect of industrial work,
these women were trained for individual tasks
% --- such as riveting component to another ---
and assigned to that task. %, decomposed from a more comprehensive task.
``Rosie the Riveter'',
an icon of \nth{20} century America who
represented empowerment and opportunity for women~\cite{honey1985creating},
was a pieceworker~\cite{davies2014origins}.

% \msb{even further backwards??? I would have expected this section to take a time slice and tell me about one group of people in detail}
% \msb{this content might go above in the paragraph about the heydey of piecework in the piecework review}

@@
